Binocular integration of chromatic and luminance signals

### 2024-07-16 manuscript.qmd
Revised in line with reviewer comments

---
title: "Binocular integration of chromatic and luminance signals"
author:
  - name: Daniel H. Baker$^{1,2,*}$
  - name: Kirralise J. Hansford$^1$
  - name: Federico G. Segala$^1$
  - name: Annie Y. Morsi$^1$
  - name: Rowan J. Huxley$^{1,3}$
  - name: Joel T. Martin$^{1,4}$
  - name: Maya Rockman$^1$
  - name: Alex R. Wade$^{1,2}$
pdf-engine: xelatex
bibliography: references.bib
csl: journal-of-vision.csl
format:
  pdf:
    keep-tex: true
#   docx
# prefer-html: true
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
---

@@

if (processdata>0){
  install_python(version="3.6")
  py_install('https://github.com/wichmann-lab/python-psignifit/zipball/master')

  # use_miniconda('r-reticulate')

@@

ticklocs <- seq(-1,1,0.5)
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-1,1), ylim=c(-1,1))  
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocs, side = 1, at=ticklocs, line=0.3)    
mtext(text = ticklocs, side = 2, at=ticklocs, line=0.2, las=1)
title(xlab="L", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="M", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='(a) L-M gratings', col.lab=rgb(0,0,0), line=1.5, cex.lab=2)

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-1,1), ylim=c(-1,1))  
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocs, side = 1, at=ticklocs, line=0.3)    
mtext(text = ticklocs, side = 2, at=ticklocs, line=0.2, las=1)
title(xlab="L+M", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="S", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='(b) S-(L+M) gratings', col.lab=rgb(0,0,0), line=1.5, cex.lab=2)

@@

  paneltitles <- c('(a) Achromatic thresholds','(b) Red/Green thresholds','(c) Blue/Yellow thresholds')
  par(mfrow=c(2,3))
  for (expt in 1:3){

@@

  paneltitles <- c('(d) Achromatic slopes','(e) Red/Green slopes','(f) Blue/Yellow slopes')

@@

  if (param==1){
    legend(0,1,c('Achromatic','Red/green','Blue/yellow','Flicker'),fill=addalpha(concols1,alpha=0.25),border=concols2,box.lwd=2,cex=0.95)
  }

@@

paneltitles <- c('(a) Achromatic grating','(b) Red/green grating','(c) Blue/yellow grating','(d) Flickering disc')
collist <- c('red','blue','orange','darkgreen')

@@

titlelist <- c('(a) Achromatic target','(b) R/G target','(c) B/Y target','(d) Achromatic densities','(e) R/G densities','(f) B/Y densities','(g) Threshold elevation','(h) Psychometric slope')

@@

  legend(-2,40,c('Target only','Achromatic mask','R/G mask','B/Y mask'), pch=21, col=c('black','black','red','cornflowerblue'),pt.cex=2,cex=1.5,pt.lwd=3,pt.bg=c('black','white','green','gold'),box.lwd=2)

@@

plotlims <- c(0.5,3.5,0.5,3.5)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocs <- seq(1,3)    # locations of tick marks on x axis
ticklabels <- c('ACH','R/G','B/Y')
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabels, side = 1, at=ticklocs,line=0.5) 
mtext(text = ticklabels, side = 2, at=ticklocs, line=0.2, las=1) 
title(xlab="Target", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Mask", col.lab=rgb(0,0,0), line=2.5, cex.lab=1.5)
title(main=titlelist[7],cex.main=2)

@@

plotlims <- c(0.5,3.5,0.5,3.5)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocs <- seq(1,3)    # locations of tick marks on x axis
ticklabels <- c('ACH','R/G','B/Y')
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabels, side = 1, at=ticklocs,line=0.5) 
mtext(text = ticklabels, side = 2, at=ticklocs, line=0.2, las=1) 
title(xlab="Target", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
title(ylab="Mask", col.lab=rgb(0,0,0), line=2.5, cex.lab=1.5)
title(main=titlelist[8],cex.main=2)

@@

    if (c[i]==2) { // red/green detection condition
    Lt = pow(TargetContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + (TargetContrast[i])) ;
    Rt = 0 ;
    Ln = 0 ;
    Rn = 0 ;

@@

    if (c[i]==3) { // blue/yellow detection condition
    Lt = pow(TargetContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4],"  + (TargetContrast[i])) ;
    Rt = 0 ;
    Ln = 0 ;
    Rn = 0 ;

@@

    if (c[i]==5) { // red/green target, achromatic mask
      Lt = pow(TargetContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + TargetContrast[i] + wRgA[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==6) { // blue/yellow target, achromatic mask
      Lt = pow(TargetContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + TargetContrast[i] + wByA[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==7) { // achromatic target, red/green mask
      Lt = pow(TargetContrast[i],",allstanparams[1,3],") / (",allstanparams[1,4]," + TargetContrast[i] + wAcR[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==8) { // red/green target, red/green mask
      Lt = pow(TargetContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + TargetContrast[i] + ",allstanparams[2,6],"*PedestalContrast[i]) ;
      Rt = pow(PedestalContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + PedestalContrast[i] + ",allstanparams[2,6],"*TargetContrast[i]) ;
      Ln = 0 ;
      Rn = pow(PedestalContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + PedestalContrast[i]) ;

@@

    if (c[i]==9) { // blue/yellow target, red/green mask
      Lt = pow(TargetContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + TargetContrast[i] + wByR[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

     if (c[i]==10) { // achromatic target, blue/yellow mask
      Lt = pow(TargetContrast[i],",allstanparams[1,3],") / (",allstanparams[1,4]," + TargetContrast[i] + wAcB[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==11) { // red/green target, blue/yellow mask
      Lt = pow(TargetContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + TargetContrast[i] + wRgB[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==12) { // blue/yellow target, blue/yellow mask
      Lt = pow(TargetContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + TargetContrast[i] + ",allstanparams[3,6],"*PedestalContrast[i]) ;
      Rt = pow(PedestalContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + PedestalContrast[i] + ",allstanparams[3,6],"*TargetContrast[i]) ;
      Ln = 0 ;
      Rn = pow(PedestalContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + PedestalContrast[i]) ;

@@

  titlelist <- c('(a) Achromatic target','(b) R/G target','(c) B/Y target')

@@

    # legend(-2,40,c('Target only','Achromatic mask','R/G mask','B/Y mask'), pch=21, col=c('black','black','red','cornflowerblue'),pt.cex=2,cex=1.5,pt.lwd=3,pt.bg=c('black','white','green','gold'),box.lwd=2)

@@

  plotlims <- c(0.5,3.5,0.5,3.5)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
  ticklocs <- seq(1,3)    # locations of tick marks on x axis
  ticklabels <- c('ACH','R/G','B/Y')
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
  axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
  mtext(text = ticklabels, side = 1, at=ticklocs,line=0.5) 
  mtext(text = ticklabels, side = 2, at=ticklocs, line=0.2, las=1) 
  title(xlab="Target", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
  title(ylab="Mask", col.lab=rgb(0,0,0), line=2.5, cex.lab=1.5)
  title(main='(d) Suppressive weights',cex.main=2)

@@

  legend(0,1,c('Target only','Achromatic mask','R/G mask','B/Y mask'), lwd=3, col=c('black','black','red','cornflowerblue'),cex=1.9,box.lwd=2)

  legend(0,1,c('Target only','Achromatic mask','R/G mask','B/Y mask'), lty=2, lwd=3, col=c('black','white','green','gold'),cex=1.9,box.lwd=2)

  legend(0.05,0.5,c('Ach target, R/G mask','Ach target, B/Y mask','R/G target, Ach mask','R/G target, B/Y mask','B/Y target, Ach mask','B/Y target, R/G mask'), lwd=3, pch=22, col=c('black','black','red','red','cornflowerblue','cornflowerblue'),pt.bg=addalpha(c('red','cornflowerblue','black','cornflowerblue','black','red'),alpha=0.5),lty=NA,pt.cex=2,cex=1.5,box.lwd=2)

@@

text(0.5,2.8,'Achromatic grating',adj=0.5,cex=1.5)
text(2,2.8,'Red/green grating',adj=0.5,cex=1.5)
text(3.5,2.8,'Blue/yellow grating',adj=0.5,cex=1.5)
text(3,1.8,'Flickering disc (4Hz)',adj=0.5,cex=1.5)

@@

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-24,30), ylim=c(-24,30))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Normalized pedestal contrast", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Normalized threshold", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main=paste0('(',letters[(3*(p-1)+2)],')  P',p,': Red/Green'),cex.main=2)
axis.break(axis=1,breakpos=-18)

@@

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-24,30), ylim=c(-24,30))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Normalized pedestal contrast", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(ylab="Normalized threshold", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main=paste0('(',letters[(3*(p-1)+3)],')  P',p,': Blue/Yellow'),cex.main=2)
axis.break(axis=1,breakpos=-18)

@@

titlelist <- c('(a)  P1: Achromatic target','(b)  P1: R/G target','(c)  P1: B/Y target','(d)  P4: Achromatic target','(e)  P4: R/G target','(f)  P4: B/Y target','(g)  P5: Achromatic target','(h)  P5: R/G target','(i)  P5: B/Y target')

@@

if (p==1 & cond==1){legend(-2,40,c('Target only','Achromatic mask','R/G mask','B/Y mask'), pch=21, col=c('black','black','red','cornflowerblue'),pt.cex=2,cex=1.5,pt.lwd=3,pt.bg=c('black','white','green','gold'),box.lwd=2)}

@@

Much progress has been made in understanding how the brain combines signals from the two eyes. However, most of this work has involved achromatic (black and white) stimuli, and it is not clear if the same processes apply in colour-sensitive pathways. In our first experiment, we measured contrast discrimination ('dipper') functions for four key ocular configurations (monocular, binocular, half-binocular and dichoptic), for achromatic, isoluminant red/green and isoluminant blue/yellow sine-wave grating stimuli. We find a similar pattern of results across stimuli, implying equivalently strong interocular suppression within each pathway. Our second experiment measured dichoptic masking within and between pathways using the method of constant stimuli. Masking was strongest within-pathway, and weakest between blue/yellow and achromatic mechanisms. Finally, we repeated the dipper experiment using temporal luminance modulations, which produced slightly weaker interocular suppression than for spatially modulated stimuli. We interpret our results in the context of a contemporary two-stage model of binocular contrast gain control, implemented here using a hierarchical Bayesian framework. Posterior distributions of the weight of interocular suppression overlapped with a value of 1 for all dipper data sets, and the model captured well the pattern of thresholds from all three experiments.

@@

The process by which the brain combines independent inputs is of fundamental importance for understanding sensory perception. Binocular vision is a useful test-case for determining the general principles involved in neural signal combination, as our brains typically combine the inputs from the left and right eyes to provide binocular single vision. In recent years our understanding has been facilitated by the development of binocular gain control models that provide a framework to interpret empirical data from multiple experimental paradigms and techniques, including psychophysics [@Meese2006], EEG [@Baker2017], fMRI [@Moradi2009] and pupillometry [@Segala2023]. However, most of this work has used achromatic (black and white) stimuli; we know comparatively little about how chromatic signals are combined binocularly, or about how signals in different ocular and chromatic channels interact. In this study we use psychophysical detection and discrimination paradigms to explore binocular interactions within and between the chromatic and achromatic pathways.

A useful framework for understanding binocular signal processing is the two-stage gain control model of binocular combination introduced by @Meese2006. This model features interocular suppression between monocular channels, followed by binocular summation. The model accounts well for the pattern of contrast discrimination ('dipper') functions for four distinct ocular configurations [see also @Georgeson2016], illustrated in @fig-exampledips. In the monocular condition, participants must discriminate between stimuli of two contrasts (a 'pedestal', and a 'pedestal plus target') that are both presented to one eye, whilst the other eye views mean luminance. Threshold is defined as the minimum target contrast required to make this judgement with 75% accuracy. The binocular condition is the same, except that the stimuli are shown to both eyes. In the half-binocular condition the pedestal is shown to both eyes, but the target increment is shown only to one eye. Finally, the dichoptic condition involves presenting the pedestal to one eye, and the target increment to the other eye.

@@

All experiments were completed by the first author (DHB) and two additional participants, who differed for each experiment. Participants had no known abnormalities of binocular or colour vision. Written informed consent was obtained before data collection began, and all procedures were approved by the ethics committee of the Department of Psychology at the University of York (ID number 2202).

>>>>

---
title: "Binocular integration of chromatic and luminance signals"
author:
  - name: Daniel H. Baker$^{1,2,*}$
  - name: Kirralise J. Hansford$^1$
  - name: Federico G. Segala$^1$
  - name: Annie Y. Morsi$^1$
  - name: Rowan J. Huxley$^{1,3}$
  - name: Joel T. Martin$^{1,4}$
  - name: Maya Rockman$^1$
  - name: Alex R. Wade$^{1,2}$
pdf-engine: xelatex
bibliography: references.bib
csl: journal-of-vision.csl
format:
  pdf:
    keep-tex: true
#   docx
# prefer-html: true
header-includes: 
  \usepackage{lineno} \linenumbers
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
---

@@

if (processdata>0){
  # install_python(version="3.6")
  # py_install('https://github.com/wichmann-lab/python-psignifit/zipball/master')

  use_miniconda('r-reticulate')

@@

ticklocs <- seq(-1,1,0.5)
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-1,1), ylim=c(-1,1))  
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocs, side = 1, at=ticklocs, line=0.3)    
mtext(text = ticklocs, side = 2, at=ticklocs, line=0.2, las=1)
title(xlab=expression(paste(Delta,"L/L")), col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab=expression(paste(Delta,"M/M")), col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='(a) L-M gratings', col.lab=rgb(0,0,0), line=1.5, cex.lab=2)

@@

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-1,1), ylim=c(-1,1))  
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocs, side = 1, at=ticklocs, line=0.3)    
mtext(text = ticklocs, side = 2, at=ticklocs, line=0.2, las=1)
title(xlab=expression(paste(Delta,"L/L + ",Delta,"M/M")), col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab=expression(paste(Delta,"S/S")), col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main='(b) S-(L+M) gratings', col.lab=rgb(0,0,0), line=1.5, cex.lab=2)

@@

  paneltitles <- c('(a) Achromatic thresholds','(b) L-M thresholds','(c) S-(L+M) thresholds')
  par(mfrow=c(2,3))
  for (expt in 1:3){

@@

  paneltitles <- c('(d) Achromatic slopes','(e) L-M slopes','(f) S-(L+M) slopes')

@@

  if (param==1){
    legend(0,1,c('Achromatic','L-M','S-(L+M)','Flicker'),fill=addalpha(concols1,alpha=0.25),border=concols2,box.lwd=2,cex=0.95)
  }

@@

paneltitles <- c('(a) Achromatic grating','(b) L-M grating','(c) S-(L+M) grating','(d) Flickering disc')
collist <- c('red','blue','orange','darkgreen')

@@

titlelist <- c('(a) Achromatic target','(b) L-M target','(c) S-(L+M) target','(d) Achromatic densities','(e) L-M densities','(f) S-(L+M) densities','(g) Threshold elevation','(h) Psychometric slope')

@@

  legend(-2,40,c('Target only','Achromatic mask','L-M mask','S-(L+M) mask'), pch=21, col=c('black','black','red','cornflowerblue'),pt.cex=2,cex=1.5,pt.lwd=3,pt.bg=c('black','white','green','gold'),box.lwd=2)

@@

plotlims <- c(0.5,3.5,0.5,3.5)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocs <- seq(1,3)    # locations of tick marks on x axis
ticklabels <- c('ACH','L-M','S-(L+M)')
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabels, side = 1, at=ticklocs,line=0.5) 
mtext(text = ticklabels, side = 2, at=ticklocs, line=0.2, las=1) 
title(xlab="Target", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)    # titles for axes
title(ylab="Mask", col.lab=rgb(0,0,0), line=3, cex.lab=1.5)
title(main=titlelist[7],cex.main=2)

@@

plotlims <- c(0.5,3.5,0.5,3.5)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
ticklocs <- seq(1,3)    # locations of tick marks on x axis
ticklabels <- c('ACH','L-M','S-(L+M)')
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabels, side = 1, at=ticklocs,line=0.5) 
mtext(text = ticklabels, side = 2, at=ticklocs, line=0.2, las=1) 
title(xlab="Target", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)    # titles for axes
title(ylab="Mask", col.lab=rgb(0,0,0), line=3, cex.lab=1.5)
title(main=titlelist[8],cex.main=2)

@@

    if (c[i]==2) { // L-M detection condition
    Lt = pow(TargetContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + (TargetContrast[i])) ;
    Rt = 0 ;
    Ln = 0 ;
    Rn = 0 ;

@@

    if (c[i]==3) { // S-(L+M) detection condition
    Lt = pow(TargetContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4],"  + (TargetContrast[i])) ;
    Rt = 0 ;
    Ln = 0 ;
    Rn = 0 ;

@@

    if (c[i]==5) { // L-M target, achromatic mask
      Lt = pow(TargetContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + TargetContrast[i] + wRgA[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==6) { // S-(L+M) target, achromatic mask
      Lt = pow(TargetContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + TargetContrast[i] + wByA[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==7) { // achromatic target, L-M mask
      Lt = pow(TargetContrast[i],",allstanparams[1,3],") / (",allstanparams[1,4]," + TargetContrast[i] + wAcR[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==8) { // L-M target, L-M mask
      Lt = pow(TargetContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + TargetContrast[i] + ",allstanparams[2,6],"*PedestalContrast[i]) ;
      Rt = pow(PedestalContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + PedestalContrast[i] + ",allstanparams[2,6],"*TargetContrast[i]) ;
      Ln = 0 ;
      Rn = pow(PedestalContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + PedestalContrast[i]) ;

@@

    if (c[i]==9) { // S-(L+M) target, L-M mask
      Lt = pow(TargetContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + TargetContrast[i] + wByR[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

     if (c[i]==10) { // achromatic target, S-(L+M) mask
      Lt = pow(TargetContrast[i],",allstanparams[1,3],") / (",allstanparams[1,4]," + TargetContrast[i] + wAcB[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==11) { // L-M target, S-(L+M) mask
      Lt = pow(TargetContrast[i],",allstanparams[2,3],") / (",allstanparams[2,4]," + TargetContrast[i] + wRgB[s[i]]*PedestalContrast[i]) ;
      Rt = 0 ;
      Ln = 0 ;
      Rn = 0 ;

@@

    if (c[i]==12) { // S-(L+M) target, S-(L+M) mask
      Lt = pow(TargetContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + TargetContrast[i] + ",allstanparams[3,6],"*PedestalContrast[i]) ;
      Rt = pow(PedestalContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + PedestalContrast[i] + ",allstanparams[3,6],"*TargetContrast[i]) ;
      Ln = 0 ;
      Rn = pow(PedestalContrast[i],",allstanparams[3,3],") / (",allstanparams[3,4]," + PedestalContrast[i]) ;

@@

  titlelist <- c('(a) Achromatic target','(b) L-M target','(c) S-(L+M) target')

@@

    # legend(-2,40,c('Target only','Achromatic mask','L-M mask','S-(L+M) mask'), pch=21, col=c('black','black','red','cornflowerblue'),pt.cex=2,cex=1.5,pt.lwd=3,pt.bg=c('black','white','green','gold'),box.lwd=2)

@@

  plotlims <- c(0.5,3.5,0.5,3.5)  # define the x and y limits of the plot (minx,maxx,miny,maxy)
  ticklocs <- seq(1,3)    # locations of tick marks on x axis
  ticklabels <- c('ACH','L-M','S-(L+M)')
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
  axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
  axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
  mtext(text = ticklabels, side = 1, at=ticklocs,line=0.5) 
  mtext(text = ticklabels, side = 2, at=ticklocs, line=0.2, las=1) 
  title(xlab="Target", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)    # titles for axes
  title(ylab="Mask", col.lab=rgb(0,0,0), line=2.5, cex.lab=1.5)
  title(main='(d) Suppressive weights',cex.main=2)

@@

  legend(0,1,c('Target only','Achromatic mask','L-M mask','S-(L+M) mask'), lwd=3, col=c('black','black','red','cornflowerblue'),cex=1.9,box.lwd=2)

  legend(0,1,c('Target only','Achromatic mask','L-M mask','S-(L+M) mask'), lty=2, lwd=3, col=c('black','white','green','gold'),cex=1.9,box.lwd=2)

  legend(0.05,0.5,c('Ach target, L-M mask','Ach target, S-(L+M) mask','L-M target, Ach mask','L-M target, S-(L+M) mask','S-(L+M) target, Ach mask','S-(L+M) target, L-M mask'), lwd=3, pch=22, col=c('black','black','red','red','cornflowerblue','cornflowerblue'),pt.bg=addalpha(c('red','cornflowerblue','black','cornflowerblue','black','red'),alpha=0.5),lty=NA,pt.cex=2,cex=1.5,box.lwd=2)

@@

text(0.5,2.8,'Achromatic grating',adj=0.5,cex=1.5)
text(2,2.8,'L-M grating',adj=0.5,cex=1.5)
text(3.5,2.8,'S-(L+M) grating',adj=0.5,cex=1.5)
text(3,1.8,'Flickering disc (4Hz)',adj=0.5,cex=1.5)

@@

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-24,30), ylim=c(-24,30))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Normalized pedestal contrast", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Normalized threshold", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main=paste0('(',letters[(3*(p-1)+2)],')  P',p,': L-M'),cex.main=2)
axis.break(axis=1,breakpos=-18)

@@

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(-24,30), ylim=c(-24,30))   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.3)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Normalized pedestal contrast", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(ylab="Normalized threshold", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main=paste0('(',letters[(3*(p-1)+3)],')  P',p,': S-(L+M)'),cex.main=2)
axis.break(axis=1,breakpos=-18)

@@

titlelist <- c('(a)  P1: Achromatic target','(b)  P1: L-M target','(c)  P1: S-(L+M) target','(d)  P4: Achromatic target','(e)  P4: L-M target','(f)  P4: S-(L+M) target','(g)  P5: Achromatic target','(h)  P5: L-M target','(i)  P5: S-(L+M) target')

@@

if (p==1 & cond==1){legend(-2,40,c('Target only','Achromatic mask','L-M mask','S-(L+M) mask'), pch=21, col=c('black','black','red','cornflowerblue'),pt.cex=2,cex=1.5,pt.lwd=3,pt.bg=c('black','white','green','gold'),box.lwd=2)}

@@

Much progress has been made in understanding how the brain combines signals from the two eyes. However, most of this work has involved achromatic (black and white) stimuli, and it is not clear if the same processes apply in colour-sensitive pathways. In our first experiment, we measured contrast discrimination ('dipper') functions for four key ocular configurations (monocular, binocular, half-binocular and dichoptic), for achromatic, isoluminant L-M and isoluminant S-(L+M) sine-wave grating stimuli. We find a similar pattern of results across stimuli, implying equivalently strong interocular suppression within each pathway. Our second experiment measured dichoptic masking within and between pathways using the method of constant stimuli. Masking was strongest within-pathway, and weakest between S-(L+M) and achromatic mechanisms. Finally, we repeated the dipper experiment using temporal luminance modulations, which produced slightly weaker interocular suppression than for spatially modulated stimuli. We interpret our results in the context of a contemporary two-stage model of binocular contrast gain control, implemented here using a hierarchical Bayesian framework. Posterior distributions of the weight of interocular suppression overlapped with a value of 1 for all dipper data sets, and the model captured well the pattern of thresholds from all three experiments.

@@

The process by which the brain combines independent inputs is of fundamental importance for understanding sensory perception [@Ernst2002; @Baker2017]. Binocular vision is a useful test-case for determining the general principles involved in neural signal combination, as our brains typically combine the inputs from the left and right eyes to provide binocular single vision [@Read2021]. In recent years our understanding has been facilitated by the development of binocular gain control models that provide a framework to interpret empirical data from multiple experimental paradigms and techniques, including psychophysics [@Meese2006], EEG [@Baker2017], fMRI [@Moradi2009] and pupillometry [@Segala2023]. However, most of this work has used achromatic (black and white) stimuli; we know comparatively little about how chromatic signals are combined binocularly, or about how signals in different ocular and chromatic channels interact. In this study we use psychophysical detection and discrimination paradigms to explore binocular interactions within and between the chromatic and achromatic pathways.

A useful framework for understanding binocular signal processing is the two-stage gain control model of binocular combination introduced by @Meese2006. This model features interocular suppression between monocular channels, followed by binocular summation. The model accounts well for the pattern of contrast discrimination ('dipper') functions for four distinct ocular configurations [see also @Georgeson2016], illustrated in @fig-exampledips. In the monocular condition, participants must discriminate between stimuli of two contrasts (a 'pedestal', and a 'pedestal plus target') that are both presented to one eye, whilst the other eye views mean luminance. Threshold is defined as the minimum target contrast required to make this judgement with 75% accuracy; this reduces at pedestal contrasts around threshold (facilitation), and increases at high pedestal contrasts (masking). The binocular condition is the same, except that the stimuli are shown to both eyes. In the half-binocular condition the pedestal is shown to both eyes, but the target increment is shown only to one eye. Finally, the dichoptic condition involves presenting the pedestal to one eye, and the target increment to the other eye.

@@

All experiments were completed by the first author (DHB) and two additional participants, who differed for each experiment. Participants had no known abnormalities of binocular or colour vision. Written informed consent was obtained before data collection began, and all procedures were approved by the ethics committee of the Department of Psychology at the University of York (ID number 2202). The study conformed to the tenets of the Declaration of Helsinki.
